---
title: "Predicting the 2029 Canadian Federal Election Popular Vote Using Post-Stratification"
author: "\\small GROUP 51: CAROL XU, ISLA CHEONG, MIA SUNG, VANESSA GABRIELA KWOK"
date: November 13, 2025
subtitle: STA304 - Assignment 2
format: pdf
editor: visual
---

```{r, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
library(openintro)
library(tidyverse)
```

# 1 Introduction

<!-- Here you should have a few paragraphs of text introducing the problem, getting the reader interested/ready for the rest of the report. -->

<!-- Introduce terminology. -->

<!-- Highlight hypotheses. -->

<!-- Optional: You can also include a description of each section of this report as a last paragraph. -->

Strong demographic and regional variations significantly impact Canadian federal elections, which complicates the estimation of voting results relying solely on survey data. This report seeks to forecast the national popular vote—the share of votes received by each party nationwide—for the approaching 2029 federal election. To achieve this, we merge individual-level survey responses from the Canadian Election Study (CES) with population figures from the 2021 Canadian Census to produce nationally representative estimates (Stephenson et al., 2022; Statistics Canada, 2023).

The ability to anticipate voting results is quite beneficial since it can reveal the larger political landscape, aiding researchers and policymakers in grasping voter perspectives through numerical data about party inclinations. Even though Canada selects governments via a first-past-the-post system (Albert, 2024), the popular vote continues to be an important measure in political discourse and scholarly studies as it indicates national backing for significant parties including the Liberal Party, Conservative Party, and New Democratic Party. In this report, popular vote denotes the percentage of all legitimate votes cast across the country for a specific political party, while post-stratification indicates the method of modifying model forecasts to align with the demographic makeup of the population.

The research question guiding this analysis is: What will the popular vote distribution look like for the Liberal, Conservative, and NDP parties in the 2029 Canadian federal election? To answer this, we apply post-stratification to the results of a regression model that estimates how voting intention varies with demographic factors. Age and province are included as predictors because trends from recent elections suggest notable generational and regional cleavages in party support.

Before conducting the statistical analysis, we proposed the following hypotheses: (1) older voters are more likely to support the Conservative Party; (2) younger voters are more likely to support the NDP; and (3) meaningful variation in voting preferences exists across provinces. The remaining sections of this report outline the data sources used, describe the modeling and post-stratification procedures, present the projected popular vote estimates, and discuss the interpretation, limitations, and future directions of this work.

# 2 Data

## 2.1 Data Description

<!-- Type here a paragraph introducing the data, its context and as much info about the data collection process that you know. -->

The analysis uses two publicly available datasets: the 2021 Canadian Election Study (CES) and the 2021 Canadian Census. The CES is a nationally distributed survey that collects information on Canadians’ political attitudes, voting intention, and demographic characteristics during federal election cycles. It utilizes sampling and weighting techniques driven by probability to ensure extensive participation from eligible voters across different areas (Stephenson et al., 2022). The 2021 Census offers comprehensive demographic figures for the Canadian populace and is conducted by Statistics Canada every five years to assess age, gender, education, and geographical distribution (Statistics Canada & Statistics Canada, 2023). Together, the CES and Census allow us to combine behavioural data (voting intention) with population counts to correct for survey non-representativeness using post-stratification.

```{r, include=FALSE}


# Here you can load in and clean the census data (you may need to do the cleaning in a separate R script). 

census_data <- read_csv("canada_census2021.csv")

# You may need additional chunks, in case you want to include some of the cleaning output.

```

<!-- Type here a summary of the cleaning process. You only need to describe cleaning that you and your group did. -->

## 2.2 Data Cleaning

Both datasets required preprocessing before analysis. First, we extracted from the CES only the variables needed for modeling: age, province, gender, education, and vote intention. Age was converted into 15 five-year categories (18–19 to 85+), provinces were standardized to Statistics Canada codes, and the three northern territories were merged into a single “North” category to match Census regional reporting. Gender responses were recoded into a binary “Woman” or “Man” classification, and education levels were collapsed into three categories: high school or less, some college, and college degree. We then generated three binary variables for respondents’ intended vote choice: Liberal, Conservative, and NDP. The Census data was cleaned using the same groupings so that each demographic combination — age × province × gender × education — aligned across the two datasets. Observations missing or outside the allowable ranges for the four variables were removed. This process ensures that predictions generated from the survey model can be applied directly to Census population counts.

```{r, include = FALSE}

# Here you can load in and clean the survey data (you may want to do the cleaning in a separate R script). 

# We can load it in:
survey_data <- read_csv("ces2021.csv")

```

```{r, include = FALSE}

#### You will need to update/clean the code below based off the variables you want to use in your poststratification.

## Note: Whatever variables were chosen from the survey data need to match what's in the census data (variable names and categories)

survey_data <- 
  survey_data %>% 
  mutate(age = cps21_age,
         prov = cps21_province,
         gender = cps21_genderid,
         edu = cps21_education,
         vote_liberal = if_else(cps21_votechoice == 1, 1, 0),
         vote_conservative = if_else(cps21_votechoice == 2, 1, 0),
         vote_ndp = if_else(cps21_votechoice == 3, 1, 0)) %>% 
  select(age, prov, gender, edu, vote_liberal, vote_conservative, vote_ndp) %>%
  
  mutate(
    age = case_when(
      age >= 18 & age <= 19 ~ 7,
      age >= 20 & age <= 24 ~ 8,
      age >= 25 & age <= 29 ~ 9,
      age >= 30 & age <= 34 ~ 10,
      age >= 35 & age <= 39 ~ 11,
      age >= 40 & age <= 44 ~ 12,
      age >= 45 & age <= 49 ~ 13,
      age >= 50 & age <= 54 ~ 14,
      age >= 55 & age <= 59 ~ 15,
      age >= 60 & age <= 64 ~ 16,
      age >= 65 & age <= 69 ~ 17,
      age >= 70 & age <= 74 ~ 18,
      age >= 75 & age <= 79 ~ 19,
      age >= 80 & age <= 84 ~ 20,
      age >= 85 ~ 21,
      TRUE ~ NA_real_
      ),
    
    prov = recode(prov,
                  '1' = 48, #AB
                  '2' = 59, #BC
                  '3' = 46, #MB
                  '4' = 13, #NB
                  '5' = 10, #NL
                  '6' = 70, #North West T --> North
                  '7' = 12, #NS
                  '8' = 70, #Nunavut --> North
                  '9' = 35, #ON
                  '10' = 11, #PEI
                  '11' = 24, #QC
                  '12' = 47, #SK
                  '13' = 70 #Yukon --> North
                  ),
    
    gender = case_when(
      gender == 1 ~ 2, #Man
      gender == 2 ~ 1, #Woman
      gender %in% c(3,4) ~ NA_real_
    ),
    
    edu = case_when(
      edu %in% c(1, 2) ~ 1, #high skl or less
      edu %in% c(3, 4, 5, 6, 7, 8) ~ 2, #completed some college
      edu %in% c(9, 10, 11) ~ 3, #College degree
      edu == 12 ~ NA_real_
    )
  ) %>%
  drop_na(age, prov, gender, edu, vote_liberal) %>%
  
  mutate(
    age = factor(age,
                          levels = 7:21,
                          labels = c("18–19", "20–24", "25–29", "30–34", "35–39",
                                     "40–44", "45–49", "50–54", "55–59", "60–64",
                                     "65–69", "70–74", "75–79", "80–84", "85+")),
    prov = factor(prov,
                        levels = c(10, 11, 12, 13, 24, 35, 46, 47, 48, 59, 70),
                        labels = c("NL", "PEI", "NS", "NB", "QC", "ON",
                                   "MB", "SK", "AB", "BC", "North")),
    gender = factor(gender,
                          levels = c(1, 2),
                          labels = c("Woman", "Man")),
    
    edu = factor(edu,
                          levels = c(1, 2, 3),
                          labels = c("High school or less",
                                     "Completed some college",
                                     "College degree"))
  )
  


census_data <- census_data %>% 
  mutate(
    age = round(agegrp),
    prov = pr,
    gender = gender,
    edu = hdgree
  ) %>%
  
  select(age, prov, gender, edu) %>%
  
  mutate(
    edu = case_when(
      edu %in% c(1, 2) ~ 1, #high skl or less
      edu %in% c(3, 4, 5, 6, 7, 8) ~ 2, #completed some college
      edu %in% c(9, 10, 11, 12, 13) ~ 3, #College degree
      TRUE ~ NA_real_
      ),
    
    age = ifelse(age < 7, NA, age)
  ) %>%
  
  drop_na(age, prov, gender, edu) %>%
  
  mutate(
    gender = factor(gender,
                          levels = c(1, 2),
                          labels = c("Woman", "Man")),
    edu = factor(edu,
                          levels = c(1, 2, 3),
                          labels = c("High school or less",
                                     "Completed some college",
                                     "College degree")),
    prov = factor(prov,
                        levels = c(10, 11, 12, 13, 24, 35, 46, 47, 48, 59, 70),
                        labels = c("NL", "PEI", "NS", "NB", "QC", "ON",
                                   "MB", "SK", "AB", "BC", "North")),
    age = factor(age,
                 levels = 7:21,
                 labels = c("18–19", "20–24", "25–29", "30–34", "35–39",
                            "40–44", "45–49", "50–54", "55–59", "60–64",
                            "65–69", "70–74", "75–79", "80–84", "85+"))
    
  )

```

```{=html}
<!--Remember, you may want to use multiple datasets here, if you do end up using multiple data sets, or merging the data, be sure to describe this in the cleaning process and be sure to discuss important aspects of all the data that you used.

Include a description of the important variables.-->

The variables used in the post-stratification model and their roles are summarized as follows: 
• Age — demographic predictor strongly linked to political preference trends 
• Province — captures well-established regional differences in Canadian voting patterns 
• Gender — included as a demographic control variable 
• Education — captures socioeconomic status differences that correlate with political behaviour 
• Vote intention (Liberal, Conservative, NDP) — outcome variables of interest
```

## 2.3 Numerical Summaries

```{r, include=FALSE}

# Use this to calculate some summary measures. 
nrow(survey_data)  # total number of respondents
names(survey_data)

gender_summary <- survey_data %>%
  count(gender) %>%
  mutate(percent = n / sum(n) * 100)

print(gender_summary)

age_summary <- survey_data %>%
  count(age) %>%
  mutate(percent = n / sum(n) * 100)

print(age_summary)

prov_summary <- survey_data %>%
  count(prov) %>%
  mutate(percent = n / sum(n) * 100)

print(prov_summary)



mean(survey_data$vote_liberal, na.rm = TRUE) #liberal vote proportion
mean(survey_data$vote_conservative, na.rm = TRUE) #conservative vote proportion
mean(survey_data$vote_ndp, na.rm = TRUE) #ndp vote proportion

```

<!--Include a description of the numerical summaries. Remember you can use `r` to use inline R code.-->

Numerical summaries help illustrate the composition of the CES sample prior to post-stratification. For example, `r nrow(survey_data)` participants remained after the cleaning procedure. The age distribution is uneven, with the largest counts in middle-aged groups and fewer respondents at the youngest (18–19) and oldest (85+) ends of the spectrum. The raw (unadjusted) proportions reporting vote intention were approximately `r round(mean(survey_data$vote_liberal)*100,1)`% for the Liberal Party, `r round(mean(survey_data$vote_conservative)*100,1)`% for the Conservative Party, and `r round(mean(survey_data$vote_ndp)*100,1)`% for the NDP, indicating a moderately competitive partisan landscape within the sample. <!-- want to edit the above paragraph later -->

```{r, echo = FALSE}

# Use this to create some plots. Should probably describe both the sample and population.
# Figure 1: Age Distribution in Survey Data
ggplot(survey_data, aes(x = age)) +
  geom_bar(fill = "skyblue", color = "black") +
  labs(title = "Age Distribution in Survey Data",
       x = "Age Group",
       y = "Count") +
  theme_minimal() +
  theme(axis.text.x=element_text(size=8, angle=30, hjust=1))

# Figure 2: Proportion Voting Liberal by Age Group
ggplot(survey_data, aes(x = age, fill = factor(vote_liberal))) +
geom_bar(position = "fill") +
scale_fill_discrete(labels=c('Not voted Liberal', 'Voted Liberal')) +
labs(y = "Proportion Voting Liberal", x = "Age Group", fill = "Vote Liberal") +
theme_minimal() +
theme(axis.text.x=element_text(size=8, angle=30, hjust=1))


# Figure 3: Proportion Voting Conservative by Age Group
ggplot(survey_data, aes(x = age, fill = factor(vote_conservative))) +
geom_bar(position = "fill") +
scale_fill_discrete(labels=c('Not voted Conservative', 'Voted Conservative')) +
labs(y = "Proportion Voting Conservative", x = "Age Group", fill = "Vote Conservative") +
theme_minimal() +
theme(axis.text.x=element_text(size=8, angle=30, hjust=1))

# Figure 4: Proportion Voting NDP by Age Group
ggplot(survey_data, aes(x = age, fill = factor(vote_ndp))) +
geom_bar(position = "fill") +
scale_fill_discrete(labels=c('Not voted NDP', 'Voted NDP')) +
labs(y = "Proportion Voting NDP", x = "Age Group", fill = "Vote NDP") +
theme_minimal() +
theme(axis.text.x=element_text(size=8, angle=30, hjust=1))




```

<!-- Include a clear description of the plot(s). I would recommend one paragraph for each plot. -->

Figure 1 displays the distribution of CES respondents across age categories. The figure shows that the majority of the CES respondents were in their late-30s to late-60s compared to young adults and seniors, highlighting the vitality of adjusting estimates based on population data. Figure 2 illustrates the proportion of each age category indicating a desire to vote Liberal. Support is lowest among the youngest respondents, increases through early and middle adulthood, and remains relatively stable across the older age groups. Figures 3 and 4 show how preference for the Conservative and NDP parties differ amongst age categories; Conservatives are most supported by older age groups, whereas the NDP party is most supported by young respondents. These illustrative visualizations encourage the addition of age and province in the regression model and emphasize the demographic gradients that post-stratification seeks to address.

All analysis, such as data cleaning, numerical summaries, and visualization, was conducted utilizing the R programming language (version 4.5.1) along with the tidyverse and ggplot2 packages for data manipulation and plotting.

# 3 Methods

<!-- Include some text introducing the methodology, maybe restating the problem/goal of this analysis. -->

To address our research question—estimating the overall popular vote in the next Canadian federal election—we implemented a two-stage analytic framework. Initially, we created a model of voting intention utilizing survey data to connect demographic traits to political inclinations. Second, we applied post-stratification to combine model predictions with census population counts, producing estimates that better reflect the national electorate rather than the (non-representative) survey sample.

## 3.1 Model Specifics

<!-- Here you can describe your regression model -->

<!-- I will (incorrectly) be using a linear regression model to model the proportion of voters who will vote for Donald Trump. This is a naive model. I will only be using age, which is recorded as a numeric variable, to model the probability of voting for Donald Trump. The simple linear regression model I am using is: -->

<!-- $$ y = \beta_0+\beta_1  x_{age} + \epsilon$$ -->

<!-- Where $y$ represents the .... $\beta_0$ represents.... -->

Voting intention is a binary outcome (support for a given party vs. not), making logistic regression an appropriate modeling choice because it models probabilities bounded between 0 and 1 and is widely used for categorical dependent variables. Logistic regression also allows us to quantify how demographic factors relate to voting behaviour while enabling prediction for population subgroups.

For the Liberal Party model, we specified: $$logit(P(Y_i=1))=\beta_0+\beta_1(age_i)+\beta_2(prov_i)$$, where $Y_i$ indicates whether respondent $i$ intends to vote Liberal (1 = yes, 0 = no), $\beta_0$ is the intercept, and $\beta_1$ and $\beta_2$ represent the effects of age and province, respectively. Parallel logistic regression models were also fit for the Conservative Party and NDP.

This model assumes: (1) the relationship between predictors and the log-odds of voting for a party is linear, (2) observations are independent, and (3) all relevant predictors are measured without systematic error. The parameters of interest are the estimated coefficients and the predicted probabilities from each model, which later enter the post-stratification step.

```{r, include=FALSE}

# Creating the Model
liberal_model <- glm(vote_liberal ~ age + prov + gender + edu, data = survey_data, family = binomial)
conservative_model <- glm(vote_conservative ~ age + prov + gender + edu, data = survey_data, family = binomial)
ndp_model <- glm(vote_ndp ~ age + prov + gender + edu, data = survey_data, family = binomial)

# Model Results (to Report in Results section)
# summary(model)
# OR
# broom::tidy(model)

### Don't show the results/output here...

```

## 3.2 Post-Stratification

<!-- Here you should explain the poststratification process -->

<!-- In order to estimate the proportion of voters..... -->

<!-- To put math/LaTeX inline just use one set of dollar signs. Example: $\hat{y}^{PS}$ -->

<!-- To put math on its own line use two sets of dollar signs: -->

<!-- $$ include.your.mathematical.model.here.if.you.have.some.math.to.show $$ -->

<!-- All analysis for this report was programmed using `R version 4.0.2` \[2\]. -->

Post-stratification is a statistical technique that adjusts model-based predictions from a survey to the demographic composition of the full population. In simple terms, it addresses the question: “If each demographic group in Canada votes similarly to our survey and according to the census proportions, what would the election outcome appear like?”

Cells were defined by all combinations of age group × province x age x education, because both variables are strongly associated with political preference and are available in both the survey and census datasets.

For each demographic cell $i$:

1.  Predict the probability of support for each party using the corresponding logistic model.
2.  Multiply that probability by the census count of individuals in the cell ($N_i$).
3.  Aggregate predictions across all cells to obtain a national estimate.

The post-stratified estimate for the Liberal party is: $$\hat{y}^{PS}=\frac{\sum_iP(Y=1|age_i,prov_i)\cdot N_i}{\sum_iN_i}$$Post-stratification is appropriate here because the survey is not perfectly representative of the Canadian electorate; without adjustment, predictions would be biased toward the survey’s demographic composition. By integrating census counts, we generate estimates that better approximate real-world voting intentions.

# 4 Results

```{r, include=FALSE}

#model <- lm(vote_liberal ~ age + prov, data=survey_data)
# ensure factor levels match EXACTLY
survey_data$age_num <- as.numeric(survey_data$age)
survey_data$prov_num <- as.numeric(survey_data$prov)
survey_data$gender_num <- as.numeric(survey_data$gender)
survey_data$edu_num <- as.numeric(survey_data$edu)

census_data$age_num <- as.numeric(census_data$age)
census_data$prov_num <- as.numeric(census_data$prov)
census_data$gender_num <- as.numeric(census_data$gender)
census_data$edu_num <- as.numeric(census_data$edu)

# Creating the Model
liberal_model <- glm(vote_liberal ~ age_num + prov_num + gender_num + edu_num, data = survey_data, family = binomial)

conservative_model <- glm(vote_conservative ~ age_num + prov_num + gender_num + edu_num, data = survey_data, family = binomial)

ndp_model <- glm(vote_ndp ~ age_num + prov_num + gender_num + edu_num, data = survey_data, family = binomial)

# Model Results (to Report in Results section)
# summary(model)
# OR
# broom::tidy(model)
```

```{r, include=FALSE}

# Here I will perform the post-stratification calculation
#census_data_counts <- census_data %>% 
#  group_by(age, prov) %>% 
#  summarise(n=n())

#census_data_counts$estimate <-
#  model %>% 
#  predict(newdata = census_data_counts, type="response")

#census_data_counts %>% 
#  ungroup(age) %>% 
#  mutate(liberal_predict_prop = estimate*n) %>%
#  summarise(liberal_predict = sum(liberal_predict_prop)/sum(n))

# Build Census Cell Counts
census_data_counts <- census_data %>%
group_by(age_num, prov_num, gender_num, edu_num) %>%
summarise(n = n(), .groups = "drop")

# Add predicted probabilities for each party using the logistic models
census_data_counts <- census_data_counts %>%
mutate(
  p_liberal = predict(liberal_model, newdata = census_data_counts, type = "response"),
  p_con = predict(conservative_model, newdata = census_data_counts, type = "response"),
  p_ndp = predict(ndp_model, newdata = census_data_counts, type = "response")
)

# Remove un-estimable cells (clean and proper)
census_data_counts <- census_data_counts %>% 
  drop_na(p_liberal, p_con, p_ndp)

# Compute post-stratified national vote share for each party
poststrat_results <- census_data_counts %>%
summarise(
  Liberal = sum(p_liberal * n) / sum(n),
  Conservative = sum(p_con * n) / sum(n),
  Ndp = sum(p_ndp * n) / sum(n)
) 


```

Here you present your results. You may want to put them into a well formatted table. Be sure that there is some text describing the results.

Note: Alternatively you can use the `knitr::kable` function to create a well formatted table from your code. See here: <https://rmarkdown.rstudio.com/lesson-7.html>.

Remember you can use `r` to use inline R code.

```{r, echo = FALSE}

# Here you can include some relevant visualizations.
library(knitr)

results_table <- poststrat_results %>%
pivot_longer(everything(),
names_to = "Party",
values_to = "PredictedShare") %>%
mutate(PredictedShare = round(PredictedShare * 100, 1))

kable(results_table,
col.names = c("Party", "Predicted Popular Vote (%)"),
caption = "Table 1. Post-Stratified Estimates of Popular Vote for Major Canadian Political Parties"
)

```

|                                      | Liberal Party | Another Party |
|--------------------------------------|---------------|---------------|
| Predicted Percentage of Popular Vote | 0.286         | ??            |
| Another Statistic?                   | ??            | ??            |

: The proportions of outcome variable of interest calculated for the following political parties: Liberal.

<!-- Include an explanation/interpretation of the visualizations. Make sure to comment on the appropriateness of the assumptions/results. -->

Table 1 presents the post-stratified estimates of the predicted national popular vote for the Liberal Party, Conservative Party, and NDP. These estimates combine the logistic regression predictions with census population totals to approximate how each party would fare if every demographic group voted in proportion to our survey results.

Based on the post-stratified estimates, the predicted popular vote shares are approximately `r results_table$PredictedShare[results_table$Party == "Liberal"]`% for the Liberal Party, `r results_table$PredictedShare[results_table$Party == "Conservative"]`% for the Conservative Party, and `r results_table$PredictedShare[results_table$Party == "NDP"]`% for the NDP. These results suggest that the race is competitive, with no single party dominating support nationally.

The pattern of predicted vote shares also aligns with expectations from historical Canadian federal elections, where the Liberal and Conservative parties typically receive the highest support, followed by the NDP. Therefore, the results appear reasonable given known voting trends and demographic patterns. However, the true uncertainty of these estimates cannot be fully quantified here because confidence intervals and standard errors were not computed; future work could incorporate bootstrapping or Bayesian post-stratification to quantify variability.

Overall, the results indicate that demographic composition plays an important role in estimating election outcomes, reinforcing the value of post-stratification when survey samples are not fully representative of the voting population.

<!-- edit the above text -->

# 5 Discussion

```{=html}
<!--Here you should give a summary of the Hypotheses, Methods and Results

Highlight Key Results.

Talk about big picture.

Comment on any Weaknesses.

Comment on Future Work/Next Steps

End with a concluding paragraph to wrap up the report.-->
```

This report aimed to predict the national popular vote in the upcoming Canadian federal election using logistic regression combined with post-stratification. Our hypotheses included: (1) older demographics are more inclined to support the Conservative Party, (2) younger demographics are more inclined to support the NDP, and (3) voting tendencies differ among provinces. To investigate these hypotheses, we first modeled voting intention for each major party using survey data and demographic predictors (age, province, gender and education). We subsequently post-stratified the resulting forecasts utilizing census population totals to derive national estimates that represent the demographic makeup of Canada.

The post-stratified results indicated that the predicted national popular vote shares are approximately `r results_table$PredictedShare[results_table$Party == "Liberal"]`% for the Liberal Party, `r results_table$PredictedShare[results_table$Party == "Conservative"]`% for the Conservative Party, and `r results_table$PredictedShare[results_table$Party == "NDP"]`% for the NDP. These findings suggest a competitive electoral landscape in which no party is projected to win a dominant share of the vote. In broad terms, the results reflect historical voting trends; the Liberal and Conservative parties remain the two major competitors, with the NDP drawing substantial—but comparatively smaller—support nationally.

The demographic patterns implied by the model are consistent with our hypotheses. The logistic regression coefficients (not shown here) indicated that increasing age was associated with higher support for the Conservative Party and lower support for the NDP, while younger respondents demonstrated a stronger likelihood of supporting the NDP. Provincial variation in predicted vote probability also supports the hypothesis that geography plays an important role in political preference. This reinforces the value of incorporating demographic variables when modeling voting intention and highlights the usefulness of post-stratification for correcting demographic imbalance in survey samples.

Even with these advantages, the analysis has multiple shortcomings. Initially, the models rely solely on age, province, gender and education as predictors, which implies that significant variables like income, urbanicity, and race/ethnicity are overlooked due to either data availability issues or inconsistencies between the survey and the census. Additionally, survey data naturally carry uncertainty, and sampling bias can remain even following post-stratification. Ultimately, we did not calculate standard errors or confidence intervals for the final popular vote estimates, meaning the uncertainty linked to the predictions cannot be formally assessed.

Future research might enhance this analysis by adding more demographic indicators, post-stratifying within finer geographic units (e.g., electoral districts), or gathering larger and more representative survey samples. Bootstrapped or Bayesian post-stratification techniques might offer uncertainty ranges for estimates, improving both interpretability and resilience. Incorporating temporal dynamics, like monthly trends prior to an election, would enable time-series forecasting instead of just a one-time estimate.

In summary, this report shows that integrating logistic regression with post-stratification yields more accurate estimates of voting intention compared to using survey data by itself. Although the results have limitations, the analytical approach uncovers significant demographic trends in political preference and offers a reproducible model for forecasting elections. Ongoing methodological improvements and more comprehensive datasets will enable progressively precise and insightful electoral predictions.

<!-- edit the discussion -->

\newpage

# 6 Generative AI Statement

```{=html}
<!--Here is where you can explain your usage of Generative AI tool(s). Be sure to reference any tools with inline citations.

Alternatively, if you did not use Generative AI, please include a brief statement outlining your workflow for completing this assignment.-->
```

Generative AI tools were used to support the organization of this report, help improve the clarity of explanations, and refine the overall quality of writing. AI assistance was also used in a limited way to enhance grammar and suggest alternative phrasing. All analysis, results, and interpretations were completed independently by the authors. While AI can improve efficiency and communication, it also has limitations. Its responses may contain inaccuracies or lack the critical judgment required for academic work. To ensure reliability, all AI-assisted content was carefully reviewed, cross-checked with course materials, and revised to reflect the authors’ own understanding.

# 7 Ethics Statement

```{=html}
<!--Explain how you ensured that your analysis is reproducible (e.g., documenting code, using proper statistical methods).

Since the CES 2021 data is publicly available, describe whether or not this the work completed in your report needs Research Ethics Board approval for the report the be made publicly available. Be sure to specifically discuss the privacy of human participants in this study.-->
```

To ensure the reproducibility of our analysis, all data cleaning, modeling, and post-stratification procedures were fully documented within the report and implemented using version-controlled R code. We used consistent statistical methods, clearly defined variable transformations, and provided enough detail for another researcher to replicate the workflow using the same publicly available datasets.

The 2021 Canadian Election Study (CES) and 2021 Census data used in this report are publicly accessible and contain no direct identifiers that could reveal the identity of individual participants. Because the data are already anonymized and released for public research purposes, the work conducted in this report does not require Research Ethics Board approval. Our analysis uses only aggregated demographic information and does not attempt to re-identify or make inferences about specific individuals. As such, participant privacy and confidentiality remain fully protected throughout the project.

# 8 Bibliography

1.  Stephenson, L. B., Harell, A., Rubenson, D., & Loewen, P. J. (2022). 2021 Canadian Election Study (CES) (Version 3.1, pp. 1615642, 59581151, 361922, 1884147, 225743, 2180672, 2057321, 59190631) \[Application/octet-stream,text/tab-separated-values,application/octet-stream,application/pdf,application/pdf,text/tab-separated-values,application/pdf,text/tab-separated-values\]. Harvard Dataverse. https://doi.org/10.7910/DVN/XBZHKC

2.  Statistics Canada & Statistics Canada. (2023). 2021 Census of Canada, Age, Household type, Gender, Disability, and Marital Status, by Census Tracts \[custom tabulation\] (Version 1.1, p. 102686142) \[Application/x-beyond2020\]. Borealis. https://doi.org/10.5683/SP3/UHEPKL

3.  Albert, N. (2024, February 11). The First-Past-the-Post Way of Voting is Better-than-the-Rest. C2C Journal. https://c2cjournal.ca/2024/02/the-first-past-the-post-way-of-voting-is-better-than-the-rest/

4.  Fourth reference in a consistent, formal citation style (e.g. APA).

5.  Fifth reference in a consistent, formal citation style (e.g. APA).

# 9 Appendix

Any additional notes/derivations that are supplementary to the report can be added in an appendix. This section will not be directly graded, but may be included for completion-sake.
